---
title: "Lab 2: Exploring Residential Sales in Ames"
author: "Team 6: Ryan Horsey, Dominic DiRe, Amelia Humphrey, Huu Phuoc Tran"
date: "`r Sys.Date()`"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(classdata)
library(ggplot2)
library(dplyr)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

## Step 1: Inspecting the Dataset

```{r inspect-data}
# Load the dataset
data(ames)

# View first rows
head(ames)

# Dimensions
dim(ames)

# Variable names and types
str(ames)

# Summary statistics
summary(ames)
```

## Step 2: Main Variable -- Sale Price

The variable of primary interest is **`Sale Price`**, the USD sale price of each residential property. It would be expected that if the sale price of a house increases or decreases, most other numerical variables in the dataset would change. For example, if one house's sale price is higher than another's, you might expect it to have a higher total living area in square feet.

## Step 3: Exploring Main Variable
```{r}
summary(ames$`Sale Price`)
```
```{r}
range(ames$`Sale Price`, na.rm = TRUE)
```
```{r}
ggplot(ames, aes(x = `Sale Price`)) +
  geom_histogram(bins = 30, fill = "blue") +
  labs(title = "sale price distribution",
       x = "Sale Price",
       y = "Count")
```
```{r}
sum(ames$`Sale Price` == 0)
```
The main thing that is odd is the sale prices that have 0, as well as some pretty extreme outliers. When we try to figure out what correlates with the sale price, there won't be any high corelations because of the outliers being so insanely high. It's also noteworthy that there are over 2000 rows with a sale price of 0. The mean of the data is an entire digit higher than the median. But the graph is very clearly unimodal and right skewed, and the presence of outliers is pretty obvious.



## Step 4: Related Variables

-- Ryan: YearBuilt

-- Nick: TotalLivingArea (sf)

-- Amelia: Lot (sf)

-- Huu: Acres
